### 2014-06-01 ACL2014/system.tex
section 3 edits

\subsection{Reusing Hash Functions}
\label{subsec:resuseHash}
In Section \ref{subsec:vlsh:cosine}, we showed that vanilla \lsh requires $L\times K$ hash functions. 
However generating hash functions is computationally expensive as 
it takes time to read all features and 
evaluate hash functions over all those features to generate a single bit. 
To minimize the number of hash functions computations, 
we use a trick from Andoni and Indyk \shortcite{Andoni08CACM} 
in which hash functions are reused to generate $L$ tables. $K$ is assumed to be even and $R \approx	\sqrt L$. 
We generate $f(q)=(h_1(q),h_2(q),\dots,h_{K/2}(q)))$ of length $k/2$. 
Next, we define $g(q)=(f_a,f_b)$, where $1\leq a < b \leq R$. This scheme generates $L= \frac{R (R -1)}{2}$. 
This scheme requires O($K  \sqrt L$) hash functions, instead of O($KL$).    
For rest of this paper, we use the above trick to generate $L$ hash tables with $K$-widith bit bucket-ids. 
\begin{table}
\centering
{
\small \addtolength{\tabcolsep}{-4.5pt}
\begin{tabular}{ll}
\hline
\hline
 \textbf{Symbol} &  \textbf{Description} \\
\hline 
$N$ & \# of query terms \\
$D$ & \# of features i.e. all clicked unique urls \\
\hline
\multirow{3}{*}{$K$} & \# of hash functions concatenated together \\ 
   & $g(q)=(h_1(q),h_2(q),\dots,h_k(q)))$ \\ 
	 & to generate the index of a table \\
\hline
 \multirow{2}{*}{$L$} & \# of tables generated independently \\ 
	  &  with $g_j(q)$ index, $\forall 1 \leq j \leq  L$  \\
\hline
$F$ & \# of bits flipped, $\forall 1 \leq j \leq  L$   \\
$\tau$ & $\tau$ threshold \\
Recall & fraction of similar candidates retrieved \\
Comparisons & Avg \# of pairwise comparisons per query  \\
\hline 
\end{tabular}
\caption{\footnotesize{Major Notations}}
}
\label{tab:notation}
\end{table}

\subsection{Multi Probe \lsh}
\label{sec:mblsh}
As we discussed in Section \ref{subsec:resuseHash} that generating hash functions can be computationally expensive.  
That is exactly one of the significant drawbacks of vanilla \lsh as it requires a large number of hash tables (large $L$) 
in order to achieve good recall in finding nearest neighbors.
Here, We describe four new variants of vanilla \lsh algorithm motivated by the research on \mblshf \cite{LvVLDB07}. 
The goal of  \mblshf is to get significantly better recall than the vanilla \lsh by using 
the same number of hash tables. The main concept behind \mblshf is to look 
up multiple buckets within a table that have a high probability of containing the nearest neighbors of a query. 
Note, eventhough vanila \lsh and \mblshf both have same number of hash tables ($L$); 
\mblshf requires more number of probes as it searches in multiple buckets within a table. 
The main advantage of searching in multiple buckets over generating more number of tables 
is that it takes more memory and time to generate more tables in pre-processing.

The original \mblshf algorithm is developed for Euclidean distance. 
If interested in details about \mblshf algorithm for Euclidean distance, the reader is referred to \cite{LvVLDB07}. 
Here, we propose four variants of \mblshf for cosine similarity:
\begin{itemize}
\item {\bf \rflipq:} The first variant is our baseline: first we compute the initial \lsh of a query q, which gives the $L$ bucket ids.  
Next, we create alternate bucket ids by taking each of the $L$ bucket ids and then creating alternate candidate buckets by flipping a 
set of coordinates randomly in the \lsh of just the query.  
\item {\bf \rflipb:} The second variant is another baseline similar to the previous one. Instead of just flipping the bits for only the query, 
here we flip bits for both the query and all the queries in the database. 
\item {\bf \dflipq:} The third variant is intelligent version of first variant. Instead of randomly flipping some coordinates, 
we devisely select a set of coordinates based on the distance of query term from the random hyperplane (hash function) 
that was used in to create this coordinate. 
The distance of query term from the random hyperplane is the absolute value which we get before applying the sign function on it (see Section \ref{subsec:vlsh:cosine}). 
\item {\bf \dflipb:} The fourth version is similar to third one, however 
here we flip bits for both the query and the queries in the database (intelligent version of second baseline). 
\end{itemize}

>>>>

\subsection{Reusing Hash Functions}
\label{subsec:resuseHash}
In Section \ref{subsec:vlsh:cosine}, we showed that vanilla \lsh requires $L\times K$ hash functions. 
However generating hash functions is computationally expensive as 
it takes time to read all features and 
evaluate hash functions over all those features to generate a single bit. 
To minimize the number of hash functions computations, 
we use a trick from Andoni and Indyk \shortcite{Andoni08CACM} 
in which hash functions are reused to generate $L$ tables. $K$ is assumed to be even and $R \approx	\sqrt L$. 
We generate $f(q)=(h_1(q),h_2(q),\dots,h_{K/2}(q)))$ of length $k/2$. 
Next, we define $g(q)=(f_a,f_b)$, where $1\leq a < b \leq R$. Using such pairing, we
can thus generate $L= \frac{R (R -1)}{2}$ hash indices. 
This scheme requires O($K  \sqrt L$) hash functions, instead of O($KL$).    
For rest of this paper, we use the above trick to generate $L$ hash tables with bucket indices of width $K$ bits.  
\begin{table}
\centering
{
\small \addtolength{\tabcolsep}{-4.5pt}
\begin{tabular}{ll}
\hline
\hline
 \textbf{Symbol} &  \textbf{Description} \\
\hline 
$N$ & \# of query terms \\
$D$ & \# of features i.e. all clicked unique urls \\
\hline
\multirow{3}{*}{$K$} & \# of hash functions concatenated together \\ 
   & $g(q)=(h_1(q),h_2(q),\dots,h_k(q)))$ \\ 
	 & to generate the index of a table \\
\hline
 \multirow{2}{*}{$L$} & \# of tables generated independently \\ 
	  &  with $g_j(q)$ index, $\forall 1 \leq j \leq  L$  \\
\hline
$F$ & \# of bits flipped, $\forall 1 \leq j \leq  L$   \\
$\tau$ & $\tau$ threshold \\
Recall & fraction of similar candidates retrieved \\
Comparisons & Avg \# of pairwise comparisons per query  \\
\hline 
\end{tabular}
\caption{\footnotesize{Major Notations}}
}
\label{tab:notation}
\end{table}

\subsection{Multi Probe \lsh}
\label{sec:mblsh}
As we discussed in Section \ref{subsec:resuseHash}, generating hash functions can be computationally expensive.
Since the memory required by the algorithm also scales linearly with $L$, the number of hash tables, it is desirable to have a small 
number of tables to reduce the memory footprint. The memory footprint of vanilla LSH is what makes it impractical for real applications. 
Here, we first describe four new variants of the vanilla \lsh algorithm motivated by the intuition in
\mblshf~\cite{LvVLDB07}. The aim of \mblshf is to obtain significantly higher recall than the vanilla \lsh while using
the same number of hash tables. In order to achieve this, the main intuition utilized in \mblshf is that
in addition to looking at the hash bucket that a test query $q$ falls in, it is also possible to look at the neigboring buckets
in order to find its near neigbour candidates. \mblshf in ~\cite{LvVLDB07} suggests exploring the neighboring buckets in order of the
Hamming distance from the bucket in which $q$ falls. The empirically show that these neighboring buckets contain the near neighbors with
very high probability. 
Note that, for the same number of hash tables \mblshf will require more number of probes since it searches for multiple buckets within a table. 
The main advantage of searching in multiple buckets over generating more number of tables 
is that it takes more memory and time to generate more tables in pre-processing.

The original \mblshf algorithm is developed for Euclidean distance, the details
are described in~\cite{LvVLDB07}. However, the Euclidean distance implementation does not immediately translate to our setting of cosine similarity. 
For example, in generating other the list of other buckets to look into,~\cite{LvVLDB07} utilizes the distance of the hash value 
to the boundary of the other bucket--- this
makes sense only when the hash value is a real number and not a $0/1$ bit, as it is for us. 
However, utilizing the same intuition, we present four variants  
of \mblshf for cosine similarity:
\begin{itemize}
\item {\bf \rflipq:} The first variant is our baseline. In this, we first compute the initial \lsh of a test query $q$, which gives the $L$ bucket ids.  
Next, we create alternate bucket ids by taking each of the $L$ bucket ids and then creating alternate candidate buckets by flipping a 
set of coordinates randomly in the \lsh of the test query $q$.  
\item {\bf \rflipb:} The second variant is another baseline similar to the previous one. Instead of just flipping the bits for only the test query, 
here we flip bits for both the test query and all the queries in the database. 
\item {\bf \dflipq:} The third variant is a more intelligent version of first variant. Instead of randomly flipping some coordinates
of the test query $q$, 
we select a set of coordinates based on the distance of $q$ from the random hyperplane (hash function) 
that was used in to create this coordinate. 
The distance of the test query $q$ from the random hyperplane is the absolute value which we get before applying the sign function on it 
(see Section \ref{subsec:vlsh:cosine}), i.e. is $ \mathrm{abs}(\alpha \cdot q)$ if the random hyperplane is $\alpha$. 
\item {\bf \dflipb:} The fourth version is similar to third one, however 
here we flip bits for both the test query and for the queries in the database (i.e. this is the intelligent version of the second baseline). 
\end{itemize}
